Manuscript_Github_Pchabaudi_boilerplate

### 2022-08-02 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Add subfam annotations to the chapl aapl heatmap

>>>>

```{r chapl_aapl_heatmap}

combined_data_pir_ht$locus <- cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
                                    cir_info$Gene.ID)]

#Remove the pirs in the 'NA' locus (i.e. not in a chapl or aapl)
# don't remove actual NAs because these are the ref genes.
combined_data_pir_locus_ht <- filter(combined_data_pir_ht,
                                     !grepl(locus, pattern = "NA \\(n" ))

combined_data_pir_locus_ht$locus <- ifelse(is.na(combined_data_pir_locus_ht$locus),
                                     'ref',
                                     combined_data_pir_locus_ht$locus)

#Change locus to factor and change levels so that AAPLs are before ChAPLs and are together.
combined_data_pir_locus_ht$locus <- factor(combined_data_pir_locus_ht$locus,
                                           levels = ordering(combined_data_pir_locus_ht$locus,
                                                             order_vector = c('AAPL',
                                                                              'ChAPL',
                                                                              'ref')))

clade_info <- getElement(cir_info,
                         'subfam')[match(rownames(combined_data_pir_locus_ht), 
                                                 cir_info$Geneid)]
clade_info[is.na(clade_info)] <- 'ref'
names(clade_info) <- rownames(combined_data_pir_locus_ht)
clade_ha <- rowAnnotation(
  'subfamily' = anno_simple(
    clade_info,
    col = annotation_col
  )
)
#Draw the clade legend for the subfamily
clade_lgd <- Legend(
  title = 'sub-family',
  at = names(annotation_col),
  legend_gp = gpar(fill = annotation_col)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_locus_ht[
      !colnames(combined_data_pir_locus_ht) %in% c('subfam', 
                                                   'locus')
    ])
  ))
```

```{r draw_locus_heatmap}

ht.list <- lapply(c('rMT', 'SBP'),
                  function(transmission){

                    #Get either the rMT or the SBP data
                    ht.df <- combined_data_pir_locus_ht[
                      ,
                      grepl(colnames(combined_data_pir_locus_ht), 
                            pattern = paste0(transmission))
                    ]

                    #Remove the transmission name from the column names
                    colnames(ht.df) <- str_remove(colnames(ht.df),
                                                  pattern = paste0(transmission, 
                                                                   '_'))

                    ht <- Heatmap(
                      log2(
                        1+ht.df
                      ),
                      cluster_rows = FALSE,
                      cluster_columns = FALSE,
                      show_row_names = FALSE,
                      name = 'log(TPM+1)', 
                      col = colorRamp2(c(0,
                                         maxcol/4,
                                         maxcol/2,
                                         3*maxcol/4,
                                         maxcol),
                                       viridis(5)),
                      na_col = 'white', 
                      border = TRUE,
                      border_gp = gpar(lwd = 0.1),
                      column_names_gp = gpar(fontsize = 10),
                      row_dend_width = unit(3, "cm"),
                      width = unit(4.5, 'cm'),
                      height = unit(12,'cm'),
                      column_title = paste(transmission),
                      row_title_rot = 0,
                      row_title_gp = gpar(fontsize = 10),
                      row_split = forcats::fct_shift(
                        relevel(
                          as.factor(combined_data_pir_locus_ht$locus),
                          ref = 'ref')
                      ),
                      heatmap_legend_param = list(
                        labels_gp = gpar(fontsize = 10),
                        title_gp = gpar(fontsize = 10,
                                        fontface = 'bold'),
                        at = c(0,
                               maxcol/4,
                               maxcol/2,
                               3*maxcol/4,
                               maxcol)
                      ),
                      #Only want the clade annotation on one heatmap not both.
                      left_annotation = if(transmission == 'rMT'){clade_ha}else{NULL}
                    )
                    return(ht)
                  })

if(global_save_var){
  pdf(file = 'plots/ChAPL-AAPL_TPMheatmap.pdf',
      height = 9,
      width = 6)
  draw(ht.list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'right',
       merge_legend = TRUE,
       annotation_legend_list = list(clade_lgd))
  dev.off()
}
